---
title: "Untitled"
author: "Carsten Ersch"
date: "21 December 2016"
output: pdf_document
---


# Statistical interference - Comparing means

Some examples taken from http://onlinestatbook.com/2/tests_of_means/ch10_exercises.html

In genral this is to show how to test for statistical differences between groups

## Question 1
The scores of a random sample of 8 students on a physics test are as follows: 60, 62, 67, 69, 70, 72, 75, and 78.

Test to see if the sample mean is significantly different from 65 at the .05 level. Report the t and p values.

```{r 1a}
    scores <- c(60,62,67,69,70,72,75,78)
    resA <- t.test(scores,paired = FALSE,var.equal = TRUE,mu=65)
```

so this seems to e as if the mean is not 65 but the mean is still within the confidence intervals.

When plotting the distribution and data below the relationship between the mean, median and the confidence intervals can be seen.

```{r}
library(ggplot2)
ggplot(data.frame(scores),aes(x=scores)) +geom_histogram(aes(y = ..density..),binwidth = 5) +geom_density()+theme_light()+
    geom_vline(xintercept =  median(scores),color="green")+
    geom_vline(xintercept =  resA$estimate,color="red")+
    geom_vline(xintercept =  resA$conf.int,color="blue")


```

```{r}
ggplot(data.frame(scores),aes(x="scores",y=scores))  +geom_violin() +geom_boxplot()+geom_point()+theme_light()+geom_rug(sides="r")+
    geom_hline(yintercept =  median(scores),color="green",show.legend = TRUE)+ 
    geom_hline(yintercept =  resA$estimate,color="red")+
    geom_hline(yintercept =  resA$conf.int,color="blue")
```



### b

The researcher realizes that she accidentally recorded the score that should have been 76 as 67. Are these corrected scores significantly different from 65 at the .05 level?


replacing the number 67 tith 76 and doing this again

```{r 1b}
    scores[3] <- 76
    t.test(scores,paired = FALSE,var.equal = FALSE,mu=65)
```




## Question 2

A (hypothetical) experiment is conducted on the effect of alcohol on perceptual motor ability. Ten subjects are each tested twice, once after having two drinks and once after having two glasses of water. The two tests were on two different days to give the alcohol a chance to wear off. Half of the subjects were given alcohol first and half were given water first. The scores of the 10 subjects are shown below. The first number for each subject is their performance in the "water" condition. Higher scores reflect better performance. Test to see if alcohol had a significant effect. Report the t and p values.

For me this is a paired t-test example

```{r}
library(reshape2)
 water <- c(16,15,11,20,19,14,13,15,14,16)
alcohol <- c(13,13,10,18,17,11,10,15,11,16)
ques2Data <- data.frame(water,alcohol)
ques2Data <- melt(ques2Data)
res2A <- t.test(water,alcohol,paired = TRUE)
```

Looking at the results from the t-test it can be seen that there is a difference if `r res2A$estimate` between the groups

I will use the same plots as before


```{r}
library(ggplot2)
library(dplyr)
ques2DataMean <- ques2Data %>% group_by(variable) %>% summarize_all(median)
ggplot(ques2Data,aes(x=value,color=variable)) +geom_histogram(aes(y = ..density..,fill=variable),bins = 10,alpha=0.5) +geom_density()+theme_light()+
    geom_vline(aes(xintercept=value,color=variable),data = ques2DataMean)


```

```{r}
ggplot(ques2Data,aes(x=variable,y=value))+geom_violin(aes(fill=variable)) +geom_boxplot(alpha=0.5)+geom_point()+theme_light()+geom_rug(sides="r")+
    geom_hline(aes(yintercept=value,color=variable),data = ques2DataMean,show.legend = TRUE)
```


## Question 3

The scores on a (hypothetical) vocabulary test of a group of 20 year olds and a group of 60 year olds are shown below. Test the mean difference for significance using the .05 level.

I will again use the t-test but this time it is not paired

```{r}
Y20 <- c(27,26,21,24,15,18,17,12,13)
Y60 <- c(26,29,29,29,27,16,20,27,NA)
ques3Data <- data.frame(Y20,Y60)
ques3Data <- melt(ques3Data)
res3A <- t.test(Y20,Y60,paired = FALSE,var.equal = FALSE)
```

Again looking at the distribution of the data

```{r}
ques3DataMean <- ques3Data %>% group_by(variable) %>% summarize_all(funs(median(., na.rm = TRUE)))
ggplot(ques3Data,aes(x=value,color=variable)) +geom_histogram(aes(y = ..density..,fill=variable),bins = 10,alpha=0.5) +geom_density()+theme_light()+
    geom_vline(aes(xintercept=value,color=variable),data = ques3DataMean)


```

```{r}
ggplot(ques3Data,aes(x=variable,y=value))+geom_violin(aes(fill=variable)) +geom_boxplot(alpha=0.5)+geom_point()+theme_light()+geom_rug(sides="r")+
    geom_hline(aes(yintercept=value,color=variable),data = ques3DataMean,show.legend = TRUE)
```



List the assumptions made in computing your answer.

- The variance is different in the two populations
- the samples were drawn from different populations, (not paired)


Additional notes from statistical inteference

- Power is a measure of how likely it is to detect a difference (between what?)
