---
title: "Regression Modelling Course Tests"
output: html_notebook
---

# week 1

## Question 1

```{r}

myFun <- function(u)
{
  x <- c(0.18, -1.54, 0.42, 0.95)
  w <- c(2, 1, 3, 1)
  return(sum((w*(x-u)^2)))
}

optimise(myFun,c(0.001,2))

```

## Question 2


```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

lm(y~ 0+x)
```

## Question 3


```{r}
data(mtcars)
 lm(mtcars$mpg ~ mtcars$wt)
```

## question 4

I guessed 0.25 --> wrong

I guessed 1 --> correct

## question 5

I guessed 0.16 --> wrong

I guessed 0.4 --> wrong

## Question 6

```{r}
x <- c(8.58, 10.46, 9.01, 9.64, 8.86)
scale(x)
```



## question 7

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)
lm(y~x)

```




## Question 8

I answered that nothing can be said from this information --> wrong

I answered it must be zero --> 


## question 9

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)

lm(x~x)
```




## question 10

I guessed Cor(Y,X) --> wrong

I guessed Var(Y)/Var(X) --> correct


# week 2


## question 1

```{r}
x <- c(0.61, 0.93, 0.83, 0.35, 0.54, 0.16, 0.91, 0.62, 0.62)
y <- c(0.67, 0.84, 0.6, 0.18, 0.85, 0.47, 1.1, 0.65, 0.36)


summary(lm(y~x))
```

I answered 0.053

## question 2

from the earlier result it should be 0.223


## question 3

```{r}
x <- mtcars$wt
y<- mtcars$mpg
mysumm <- lm(y ~ x)
newD <- data.frame(x = mean(x))
predict(mysumm,newdata = newD,interval = "confidence")
```

I answered 18.99




## question 4

I answered The estimated expected change in mpg per 1,000 lb increase in weight.

## question 5

```{r}
newD <- data.frame(x = mean(3))
predict(mysumm,newdata = newD,interval = "confidence",level = 0.95)
```

None of these numbers is available as an answer. I guessed 27.57 as the others are too low

## question 6

```{r}
summary(mysumm)
```

I guessed -9 --> wrong
I guessed -6.486 --> wrong

## question 7

I guessed multiplied by 100 --> correct

## question 8

I guessed The new slope would be 
c
β
^
1

this was wrong

I then guessed The new slope would be 
β
^
1
+c

this was wrong


## question 9

I guessed 0.75 -> wrong

I guessed 0.25 -> correct

## question 10

I guessed If an intercept is included, then they will sum to 0. -> correct




# Week 3



## question 1

```{r}
data("mtcars")
summary(lm(mpg ~ wt + factor(cyl), data = mtcars))


```

I answered -6.071 which is correct

## question 2


```{r}
data("mtcars")
require(lme4)
#lmList(mpg ~ wt | cyl, data = mtcars)
summary(lm(mpg ~ wt + factor(cyl), data = mtcars))$coeff
summary(lm(mpg ~ factor(cyl), data = mtcars))$coeff
```

I answered that holding weight constant cylinder has more of an effect which is wrong

I answered that it has less of an effect this time, even though I don thtink so and this is correct


## question 3


```{r}
fit1 <- lm(mpg ~ wt + factor(cyl), data = mtcars)
fit2 <- lm(mpg ~ wt * factor(cyl), data = mtcars)
anova(fit1,fit2)
```

I answered wrongly

The P-value is larger than 0.05. So, according to our criterion, we would fail to reject, which suggests that the interaction terms is necessary.

I answered wrongly

The P-value is small (less than 0.05). So, according to our criterion, we reject, which suggests that the interaction term is not necessary. 

I answered wrongly

The P-value is small (less than 0.05). Thus it is surely true that there is no interaction term in the true model.

I answered this time the answer that I actually wanted from the bgeinning but never found

The P-value is larger than 0.05. So, according to our criterion, we would fail to reject, which suggests that the interaction terms may not be necessary.

## question 4

```{r}
summary(lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars))
```

I answered wrongly
The estimated expected change in MPG per half ton increase in weight for for a specific number of cylinders (4, 6, 8).

I answered wrongly 
The estimated expected change in MPG per half ton increase in weight.

I answered wrongly
The estimated expected change in MPG per half ton increase in weight for the average number of cylinders.

I answered

The estimated expected change in MPG per half ton increase in weight for for a specific number of cylinders (4, 6, 8).


## question 5

```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
hatvalues(lm(y~x))
```

I answered 0.9946 which is correct

## question 6

```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
dfbeta(lm(y~x))
```

I answered 0.673 because the value I think it is is not there which was wrong

Then I answered -0.378 which is wrong

Then I answered -.00134 which is wrong

I answered -134


## question 7

I answered correctly

It is possible for the coefficient to reverse sign after adjustment. For example, it can be strongly significant and positive before adjustment and strongly significant and negative after adjustment.



# week 4

## question 1

```{r}
library(MASS)
shuttle <- shuttle
shuttle$use1 <- ifelse(shuttle$use %in% "auto",1,0)
exp(summary(glm(use ~ wind, data = shuttle, family="binomial" ))$coeff)

```

I answered 0.031 which is wrong

I answered 0.968 which is correct

## question 2


```{r}
exp(summary(glm(use1 ~ wind + magn, data = shuttle, family="binomial" ))$coeff)
1/1.032
```

I guessed 0.684 which is wrong

I answered 1.00 which is wrong

I answered 0.969

## question 3




```{r}
summary(glm(use1 ~ wind, data = shuttle, family="binomial" ))$coeff
summary(glm((1-use1) ~ wind, data = shuttle, family="binomial" ))$coeff
```



I answered wrongly

The intercept changes sign, but the other coefficients don't.

I answered correctly

The coefficients reverse their signs.

## question 4

```{r}
require(datasets)
data("InsectSprays")

exp(summary(glm(count ~ spray, data = InsectSprays, family="poisson" ))$coeff)
1/1.05
```


I answered -0.056 which is wrong

I answered 0.9457 which is correct

## question 5

I guessed wrongly

The coefficient is subtracted by log(10).

I guessed wrongly

The coefficient estimate is divided by 10.

I guessed

The coefficient estimate is multiplied by 10.

## question 6


I guessed 2.037 which is wrong

I guessed 1.013 which is correct











